Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-06-04 article-models-tt.tex
reorganized first section

\subsubsection{Extension}
\label{extension}
The operation of extension has its place right at the core of the theory of
contexts, families and terms since it enables one to consider families over
families and higher level families alike.

We will need the extension operation at the level of contexts and families 
over them and at the level of families and families over them. 
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jctx{\ctxext{\Gamma}{A}}
    }
& & \inference
    { \jctxeq{\Gamma}{\Delta}
      \jfameq{\Gamma}{A}{B}
      }
    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
      }
    \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfam{\Gamma}{\ctxext{A}{P}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{B} 
      \jfameq{{\Gamma}{A}}{P}{Q}
      }
    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
      }
\end{align}
The extension operation may also be defined to act on families over families
and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
action of context extension on families and terms.

@@

@@

Note that we can also weaken by the empty family over $\Gamma$ itself.
Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
their terms, families over those families and terms of those unchanged:
\label{comp-0w}\begin{align}
& \inference
  { \jfam{\Gamma}{B}
    }
  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
    }
  \label{comp-0w-c}
  \\
& \inference
  { \jfam{{\Gamma}{B}}{Q}
    }
  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
    }
  \label{comp-0w-f}
  \\
& \inference
  { \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
    }
  \label{comp-0w-t}
\end{align}

@@

\label{comp-0s}
The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
leaves everything unchanged.
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
    }
  \label{comp-0s-c}
  \\
%& \inference
%  { \jterm{\Gamma}{A}{x}
%    }
%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
%    }
%  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
    }
  \label{comp-0s-f}
  \\
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
    }.
  \label{comp-0s-t}
\end{align}

@@

@@

\subsection{The judgmental equalities of the theory of contexts, families and
terms}
In this section we extend the judgmental equality relation in such a way that
the empty context, extension, weakening, substitution and the identity terms
all become compatible with each other. Since we have five operations, there
are going to be $25$ finite sets of rules explaining how each operation is
compatible with each of the operations.

@@

The table below indicates where
we can find the axioms of each of the combinations of operators. The reference
at the row of operation $X$ and the column of operation $Y$ refers to the
axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
applied after $X$. The empty context has no arguments, whereas extension,
and weakening have two arguments which are both of the family-kind, substitution
has two arguments of which one is a term and the other is a family and
idenitity terms have one argument, a family.
\begin{center}
\begin{tabular}{r|rrrr}
& ext & wk & subst & id\\
\hline
extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
\end{tabular}
\end{center}
Apart from references to sections where axioms are given, we also see checkmarks,
crosses, marked references and duplicated references in this table. 
The checkmarks indicate that the desired property
can be proved and need not be axiomatized. To clarify those:
\begin{itemize}
\item The identity term of the empty context is the unique term of the empty
family. Recall that we already have axiomatized in \autoref{empty} that every
term of the empty family is judgmentally equal to $\emptytm$.
\item The identity term of an extension can be explained by means of the notion
of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
using only identity terms and substitution. With extension on terms we can
combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
pair consisting of $x$ and $u$. Since it is a defined notion, we will
be able to derive lots of properties for extension of terms. One of them explains
how we can see the identity term of an extension as the extension of two certain
terms, see \autoref{comp-ie}.
\item Using extension on terms we can also make sense of `fibers at an extension'
because that would exactly be the fiber at an extended term. We will see in
\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
\end{itemize}
The spots with a \xmark simply make no sense 
{\color{red}(or do they, when the term model
is better understood (by me)?)}. The marked references point to
axioms which \emph{do} have a variant for terms, but those require the notion
of extension on terms, can be proved and therefore need not be included in our
list of compatibility axioms. In each of these sections, we give a forward
reference to the lemma in which this variant is stated and proved.

\subsubsection{Extensions of extensions are extensions by extensions}
\label{comp-ee}
The inference rules asserting that extension is compatible with itself assert
that contexts are unstructured lists of type declarations. This rule is
unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
\begin{align}
%& \inference
%  { \jfam{\Gamma}{A}
%    \jfam{{\Gamma}{A}}{P}
%    }
%  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
%    }
%  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
    }
  \label{comp-ee-c}
\end{align}
Note that we may take the point of view that extension distributes over
extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
simply $Q$ since extension acts trivially at the level of families (and terms).

\subsubsection{Weakenings of extensions}
\label{comp-we}
The following rules assert that when an extended family is weakened, the
weakening distributes over the extension factors.
\begin{align}\label{comp-we-f}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jfameq
      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{\ctxext{Q}{R}}}
      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
    }
\end{align}
When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
$B$, this looks already like form of type theoretic choice. It is weaker in that
it is not stated with function types, yet it is stronger in that it states a
judgmental equality between two families.

\subsubsection{Fibers of extensions}
\label{comp-se}
The following inference rule asserts that if we take the fiber of an extended
family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
distributes over the factors of the extension.
\begin{align}\label{comp-se-f}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\subst{x}{P}}}
      {\subst{x}{\ctxext{Q}{R}}}
      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
    }
\end{align}
There is also a version of this statement in which extended terms are considered.
This variant is stated and proved in \autoref{comp-se-t}.

\subsubsection{Weakening of weakenings}
When a weakened family is weakened even further, the outer weakening distributes
over the `weakening factors'.
\label{comp-ww}
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    \jfam{{\Gamma}{C}}{R}
    }
  { \jfameq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
      {\ctxwk{A}{{B}{R}}}
      {\ctxwk{{A}{B}}{{A}{R}}}
    }
  \label{comp-ww-f}\\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    \jterm{{\Gamma}{C}}{R}{t}
    }
  { \jtermeq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
      {\ctxwk{{A}{B}}{{A}{R}}}
      {\ctxwk{A}{{B}{t}}}
      {\ctxwk{{A}{B}}{{A}{t}}}
    }
  \label{comp-ww-t}
\end{align}

\subsubsection{Fibers of weakenings}\label{comp-sw}
Substitutions distribute over weakenings:
\begin{align}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    \jfam{{{\Gamma}{A}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
      {\subst{x}{\ctxwk{P}{R}}}
      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
    }
  \label{comp-sw-f}\\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
    }
  { \jtermeq
      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
      {\subst{x}{\ctxwk{P}{R}}}
      {\subst{x}{\ctxwk{P}{h}}}
      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
    }
  \label{comp-sw-t}
\end{align}

\subsubsection{Fibers of fibers}
\label{comp-ss}

\subsubsection{Weakenings of fibers}
\label{comp-ws}
We already have rules for the compatibility of substitution with weakening, but
we still need the rules the other way around, asserting that there is a 
judgmental equality $\ctxwk{A}{\subst{y}{Q}}\jdeq\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}$
together with all its variants.

\subsubsection{The identity term of a weakened family}
\label{comp-wi}
The identity term of a weakened family is the weakened identity term:
\begin{equation}
\inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \unfoldall{\jhomeq
      {{\Gamma}{A}}
      {\ctxwk{A}{B}}
      {\ctxwk{A}{B}}
      {\ctxwk{A}{\idtm{B}}}
      {\idtm{\ctxwk{A}{B}}}
    }}
  \label{comp-wi-t}
\end{equation}

\subsubsection{The identity term of a substituted family}
\label{comp-si}
The identity term of a substituted family is the substitution of the identity term
\begin{equation}
\inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    }
  { \unfoldall{\jhomeq
      {\Gamma}
      {\subst{x}{P}}
      {\subst{x}{P}}
      {\subst{x}{\idtm{P}}}
      {\idtm{\subst{x}{P}}}
    }}
  \label{comp-si-t}
\end{equation}

>>>>

\subsubsection{Extension}
\label{extension}
The operations of extension enable one to consider families over
families and higher level families alike. We will need both a context extension
operation and a family extension operation.

The inference rules introducing context and family extension are as follows:
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jctx{\ctxext{\Gamma}{A}}
    }
& & \inference
    { \jctxeq{\Gamma}{\Delta}
      \jfameq{\Gamma}{A}{B}
      }
    { \jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}
      }
    \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfam{\Gamma}{\ctxext{A}{P}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{B} 
      \jfameq{{\Gamma}{A}}{P}{Q}
      }
    { \jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}
      }
\end{align}
The extension operation may also be defined to act on families over families
and terms thereof. For instance, When $\jfam{{\Gamma}{A}}{P}$ is a family,
we would get a family $\jfam{{\Gamma}{A}}{{\Gamma}{P}}$ and when
$\jterm{{\Gamma}{A}}{P}{f}$ is a term, we would get a term
$\jterm{{\Gamma}{A}}{{\Gamma}{P}}{\ctxext{\Gamma}{f}}$. Then there would be
axioms stating the judgmental equalities $\ctxext{\Gamma}{P}\jdeq P$ and
$\ctxext{\Gamma}{f}\jdeq f$. Thus, it is of no actual use to introduce the
action of context extension on families and terms.

@@

\paragraph{Associativity of extension}
\label{comp-ee}
The inference rules asserting that extension is compatible with itself assert
that contexts are unstructured lists of type declarations. This rule is
unavoidable if we want that for a family $A$ in context $\Gamma$, a family over
$A$ is the same thing as a family over $\ctxext{\Gamma}{A}$. 
\begin{align}
%& \inference
%  { \jfam{\Gamma}{A}
%    \jfam{{\Gamma}{A}}{P}
%    }
%  { \jctxeq{\ctxext{{\Gamma}{A}}{P}}{\ctxext{\Gamma}{{A}{P}}}
%    }
%  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq{\Gamma}{\ctxext{A}{{P}{Q}}}{\ctxext{{A}{P}}{Q}}
    }
  \label{comp-ee-c}
\end{align}
Note that we may take the point of view that extension distributes over
extensions. If we want to extend an extended family $\ctxext{P}{Q}$ over
$\ctxext{\Gamma}{A}$ by the family $A$, the result is the extension of (1)
the extension of $P$ by $A$ and (2) the extension of $Q$ by $A$. The latter is
simply $Q$ since extension acts trivially at the level of families (and terms).

@@

\paragraph{Weakenings of extensions}
\label{comp-we}
The following rules assert that when an extended family is weakened, the
weakening distributes over the extension factors.
\begin{align}\label{comp-we-f}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jfameq
      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{\ctxext{Q}{R}}}
      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
    }
\end{align}
When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
$B$, this looks already like form of type theoretic choice. It is weaker in that
it is not stated with function types, yet it is stronger in that it states a
judgmental equality between two families.

\paragraph{Weakening of weakenings}
When a weakened family is weakened even further, the outer weakening distributes
over the `weakening factors'.
\label{comp-ww}
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    \jfam{{\Gamma}{C}}{R}
    }
  { \jfameq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
      {\ctxwk{A}{{B}{R}}}
      {\ctxwk{{A}{B}}{{A}{R}}}
    }
  \label{comp-ww-f}\\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    \jterm{{\Gamma}{C}}{R}{t}
    }
  { \jtermeq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
      {\ctxwk{{A}{B}}{{A}{R}}}
      {\ctxwk{A}{{B}{t}}}
      {\ctxwk{{A}{B}}{{A}{t}}}
    }
  \label{comp-ww-t}
\end{align}

@@

@@

\paragraph{Fibers of extensions}
\label{comp-se}
The following inference rule asserts that if we take the fiber of an extended
family at a term $x$ of $A$ in context $\Gamma$, the substitution by $x$
distributes over the factors of the extension.
\begin{align}\label{comp-se-f}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\subst{x}{P}}}
      {\subst{x}{\ctxext{Q}{R}}}
      {\ctxext{\subst{x}{Q}}{\subst{x}{R}}}
    }
\end{align}
There is also a version of this statement in which extended terms are considered.
This variant is stated and proved in \autoref{comp-se-t}.

\paragraph{Fibers of weakenings}\label{comp-sw}
Substitutions distribute over weakenings:
\begin{align}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    \jfam{{{\Gamma}{A}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
      {\subst{x}{\ctxwk{P}{R}}}
      {\ctxwk{\subst{x}{P}}{\subst{x}{R}}}
    }
  \label{comp-sw-f}\\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    \jterm{{{\Gamma}{A}}{Q}}{R}{h}
    }
  { \jtermeq
      {\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\ctxwk{P}{Q}}}}
      {\subst{x}{\ctxwk{P}{R}}}
      {\subst{x}{\ctxwk{P}{h}}}
      {\ctxwk{\subst{x}{P}}{\subst{x}{h}}}
    }
  \label{comp-sw-t}
\end{align}

\paragraph{Fibers of fibers}
\label{comp-ss}

\paragraph{Weakenings of fibers}
\label{comp-ws}
We already have rules for the compatibility of substitution with weakening, but
we still need the rules the other way around, asserting that there is a 
judgmental equality $\ctxwk{A}{\subst{y}{Q}}\jdeq\subst{\ctxwk{A}{y}}{\ctxwk{A}{Q}}$
together with all its variants.

@@

\paragraph{The identity term of an extended family}
The fact that identity terms are compatible with extensions is derived in
\autoref{comp-ie}.

\paragraph{The identity term of a weakened family}
\label{comp-wi}
The identity term of a weakened family is the weakened identity term:
\begin{equation}
\inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \unfoldall{\jhomeq
      {{\Gamma}{A}}
      {\ctxwk{A}{B}}
      {\ctxwk{A}{B}}
      {\ctxwk{A}{\idtm{B}}}
      {\idtm{\ctxwk{A}{B}}}
    }}
  \label{comp-wi-t}
\end{equation}

\paragraph{The identity term of a substituted family}
\label{comp-si}
The identity term of a substituted family is the substitution of the identity term
\begin{equation}
\inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    }
  { \unfoldall{\jhomeq
      {\Gamma}
      {\subst{x}{P}}
      {\subst{x}{P}}
      {\subst{x}{\idtm{P}}}
      {\idtm{\subst{x}{P}}}
    }}
  \label{comp-si-t}
\end{equation}

\subsection{Further judgmental equalities of the theory of contexts, families and
terms}
We have already seen that empty families, extension, weakening, substitution
and the identity terms are compatible with each other. There are, however, a
few more judgmental equalities explaining what happens when
\begin{enumerate}
\item something is weakend by the empty family. This is used to consider a term
      $x$ of $A$ in context $\Gamma$ as a morphism from $\emptyf$ to $A$ in
      context $\Gamma$, see \autoref{morphisms};
\item something is substituted by the term of the empty family;
\item something is weakened by an extended family. This is Currying for weakening;
\item something is first weakened by a family $A$ in context $\Gamma$ and then
      substituted by a term of $A$. This is used to show that composition is
      compatible with itself (i.e. that composition is associative);
\item the identity terms are joining the party. Identity terms are empty contexts
      in certain models, so they should behave accordingly.
\end{enumerate}
In other words, the rules seem to not fit in the scheme that all operations are
compatible with each other, but in fact each of these rules plays a role in
completing the scheme (work out further).

\begin{comment}
In this section we extend the judgmental equality relation in such a way that
the empty context, extension, weakening, substitution and the identity terms
all become compatible with each other. Since we have five operations, there
are going to be $25$ finite sets of rules explaining how each operation is
compatible with each of the operations.

@@

The table below indicates where
we can find the axioms of each of the combinations of operators. The reference
at the row of operation $X$ and the column of operation $Y$ refers to the
axiom which explains $Y$ distributes over the arguments of $X$ when $Y$ is
applied after $X$. The empty context has no arguments, whereas extension,
and weakening have two arguments which are both of the family-kind, substitution
has two arguments of which one is a term and the other is a family and
idenitity terms have one argument, a family.
\begin{center}
\begin{tabular}{r|rrrr}
& ext & wk & subst & id\\
\hline
extension & \autoref{comp-ee}* &  & \cmark & \cmark \\
weakening & \autoref{comp-we}* & \autoref{comp-ww} & \autoref{comp-sw} & \autoref{comp-wi} \\
substitution & \autoref{comp-se}* & \autoref{comp-ws} & \autoref{comp-ss} & \autoref{comp-si} \\
identity terms & \cmark & \autoref{comp-wi} & \autoref{comp-si} & \xmark
\end{tabular}
\end{center}
Apart from references to sections where axioms are given, we also see checkmarks,
crosses, marked references and duplicated references in this table. 
The checkmarks indicate that the desired property
can be proved and need not be axiomatized. To clarify those:
\begin{itemize}
\item The identity term of the empty context is the unique term of the empty
family. Recall that we already have axiomatized in \autoref{empty} that every
term of the empty family is judgmentally equal to $\emptytm$.
\item The identity term of an extension can be explained by means of the notion
of \emph{extension on terms}. This notion is defined in \autoref{extension-on-terms}
using only identity terms and substitution. With extension on terms we can
combine a term $\jterm{\Gamma}{A}{x}$ and a term $\jterm{\Gamma}{\subst{x}{P}}{u}$
into a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$, which may be seen as the
pair consisting of $x$ and $u$. Since it is a defined notion, we will
be able to derive lots of properties for extension of terms. One of them explains
how we can see the identity term of an extension as the extension of two certain
terms, see \autoref{comp-ie}.
\item Using extension on terms we can also make sense of `fibers at an extension'
because that would exactly be the fiber at an extended term. We will see in
\autoref{comp-es} that, for instance, $\subst{\tmext{x}{u}}{Q}\jdeq\subst{u}{{x}{Q}}$
for any family $\jfam{{{\Gamma}{A}}{P}}{Q}$ and any two terms $\jterm{\Gamma}{A}{x}$
and $\jterm{\Gamma}{\subst{x}{P}}{u}$. That's a form of currying.
\end{itemize}
The spots with a \xmark simply make no sense 
{\color{red}(or do they, when the term model
is better understood (by me)?)}. The marked references point to
axioms which \emph{do} have a variant for terms, but those require the notion
of extension on terms, can be proved and therefore need not be included in our
list of compatibility axioms. In each of these sections, we give a forward
reference to the lemma in which this variant is stated and proved.
\end{comment}

\subsubsection{Weakening by the empty family}
Note that we can also weaken by the empty family over $\Gamma$.
Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
their terms, families over those families and terms of those unchanged:
\label{comp-0w}
\begin{align}
& \inference
  { \jfam{\Gamma}{B}
    }
  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
    }
  \label{comp-0w-c}
  \\
& \inference
  { \jfam{{\Gamma}{B}}{Q}
    }
  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
    }
  \label{comp-0w-f}
  \\
& \inference
  { \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
    }
  \label{comp-0w-t}
\end{align}

\subsubsection{Substituting by the term of the empty family}
\label{comp-0s}
The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
leaves everything unchanged.
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
    }
  \label{comp-0s-c}
  \\
%& \inference
%  { \jterm{\Gamma}{A}{x}
%    }
%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
%    }
%  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
    }
  \label{comp-0s-f}
  \\
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
    }.
  \label{comp-0s-t}
\end{align}
